Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-12-29 supplement.tex
section of bootstrap fits

\subsection{Likelihoods, uncertainty, and confidence intervals}

@@

For the best fit parameters, we computed confidence intervals using the same
set of bootstrapped samples and Godambe Information approach
\citep{Coffman2016-yq}, which corrects composite likelihood estimates of
confidence intervals to account for nonindependence in the data, including
linkage between loci and nonindependence of recombination bins.

\aprcomment{To ensure that our confidence intervals computed using the approach
    from \citet{Coffman2016-yq} do not underestimate true CIs, we also
    performed block bootstrap resampling of our data, 200 times, and refit each
    model to the bootstrapped datasets. From these, we evaluated the
distribution of inferred parameters, which should be similar to or narrower
than the CIs estimated from the GIM approach. See
Section~\ref{sec:bootstrap-CIs}.}

@@

\aprcomment{Here, point to tables for CIs, or SEs from Godambe vs reinferred
models from bootstrapped data. See Table~\ref{tab:supp-confidence-intervals}}

@@

%\begin{table}[ht]
%    \caption{
%        \label{tab:supp-OOA-SFS}
%        \textbf{OOA parameters using LD and SFS statistics.}
%        \aprcomment{To justify 60ka}
%    }
%    \centering
%    \begin{tabular}[t]{rcc}
%        \toprule
%        Parameter & SFS from synonymous variation & LD from intergenic variation \\
%        \midrule
%        $N_e$ &  & \\
%        $N_{AF}$ &  & \\
%        $N_{B}$ &  & \\
%        $N_{EU0}$ &  & \\
%        $N_{EUF}$ &   & \\
%        $N_{AS0}$ &  & \\
%        $N_{ASF}$ &  & \\
%        $m_{AF-B}$  & & \\
%        $m_{AF-EU}$  & & \\
%        $m_{AF-AS}$  & & \\
%        $m_{EU-AS}$  & & \\
%        $T_{EXP}$ &  & \\
%        $T_{OOA}$ &  & \\
%        $T_{EU-AS}$  & & \\
%        \bottomrule
%    \end{tabular}
%\end{table}

>>>>

\subsection{Likelihoods, uncertainty, and confidence intervals}
\label{sec:likelihoods-uncertainty}

@@

For the best fit parameters, we computed $95\%$ confidence intervals using the
same set of bootstrapped samples and Godambe Information approach
\citep{Coffman2016-yq}, which corrects composite likelihood estimates of
confidence intervals to account for nonindependence in the data, including
linkage between loci and nonindependence of recombination bins.  To ensure that
our confidence intervals computed using the approach from
\citet{Coffman2016-yq} do not underestimate true CIs, we also performed block
bootstrap resampling of our data, 200 times, and refit each model to the
bootstrapped datasets. From these, we evaluated the distribution of inferred
parameters, which should be similar to or narrower than the CIs estimated from
the GIM approach (Section~\ref{sec:bootstrap-CIs}).

@@

We created 200 block-bootstrap replicate datasets and refit our four reported
models to each replicate (Section~\ref{sec:likelihoods-uncertainty}). From
these refit models, we estimated the empirical uncertainty by taking the
standard deviation of the distribution of each refit parameter. We compared
these to our estimated standard errors from the Godambe Information Matrix
(GIM) approach to check that the confidence intervals from the GIM approach do
not underestimate true uncertainty (Table~\ref{tab:supp-confidence-intervals}).

The GIM-computed standard errors are generally close to those from the fits to
the bootstrap data, and some are larger (roughly 2--5x) than those from the
bootstrap fits. A possible explanation for this discrepancy is that some
parameters of our models come close to the lower and upper bounds allowed for
those parameters (e.g., small recent population size in the Nama, a bottleneck
in one of the stems). In such cases, the distribution of those reinferred
parameters and other correlated parameters in the model may be narrower than
the true uncertainties in those parameters. Model constraints and parameter
bounds in optimization may also lead to parameters fits that are away from
their true optimal value. In this case, the GIM-estimated confidence intervals
should be larger and contain what would be found from an unconstrained fit.

To demonstrate this effect, we simulated data under a simple
isolation-with-migration model, in which population sizes are allowed to vary,
and one population has a severe bottleneck in the most recent 1000 years
(Figure~\ref{fig:supp-uncerts-toy}). We simulated 200 1Mb regions with
$u=r=10^{-8}$, sampling 10 diploid individuals from each present-day population
and computed LD statistics. From these, we inferred the best fit parameters
without imposing bounds on the parameters, and we also reinferred the model
imposing a lower bound of 300 on population sizes (the bottleneck size). From
our fits, we computed confidence intervals using the Fisher information matrix
(FIM) and GIM approaches, in both the unbounded and bounded fits. We also
reinferred parameters over 200 bootstrapped replicate datasets to compare
empirical uncertainties to those estimated CIs.

As expected the FIM CIs underestimate the true uncertainty and the GIM CIs
closely match empirical uncertainty from the bootstrap fits. In the fit that
imposed a lower bound on population sizes, some GIM-estimated standard errors
were larger than those from the unbounded fits.  However, the fit parameters
were slightly shifted from the unbounded fits, due to hitting the lower bound
of the bottleneck size. In these cases, the estimated CIs were larger but
contained the true CIs from the unbounded GIM and bootstrap fits. For parameters
that were not strongly affected by hitting the bounds, the CIs matched closely
(Figure~\ref{fig:supp-uncerts-toy}).

\begin{figure}[htb]
    \centering
    \includegraphics{./figures/supp-uncerts-toy.pdf}
    \caption{
        \textbf{Evaluation of GIM confidence intervals using moments-LD.}
        The GIM approach accurately estimates confidence intervals, accounting
        for non-independence between regions of linked loci. GIM-estimated CIs
        from the constrained fit are large enough to contain the true
        uncertainty. Vertical dashed lines indicate simulated parameter values.
    }
    \label{fig:supp-uncerts-toy}
\end{figure}

@@
